Containers and R---the Rockerverse and beyond

### 2020-01-16 manuscript.Rmd
small shortening of bioconductor section

_Bioconductor_ ([https://bioconductor.org/](https://bioconductor.org/)) is an open source, open development project for the analysis and comprehension of genomic data \citep{gentleman_bioconductor_2004}.
The project consists of 1741 R software packages as of August 15th 2019, as well as packages containing annotation or experiment data.
_Bioconductor_has a semiannual release cycle, each release is associated with a particular version of R.
Docker images allow availability of current and past versions of _Bioconductor_ for convenience and reproducibility.
_Bioconductor_ 'base' docker images are built on top of `rocker/r-ver` and `rocker/rstudio`.
_Bioconductor_ installs packages based on the R version, and therefore uses `rocker/rstudio` and `rocker/r-ver` version tagging.
_Bioconductor_ selects the desired version of R from Rocker, adds the BiocManager CRAN package for installing appropriate versions of _Bioconductor_ packages, and creates a _Bioconductor_ docker image with an informative tag (R\_version\_Bioc\_version).
The images are summarized on the _Bioconductor_ web site (https://bioconductor.org/help/docker/), maintained on GitHub ([https://github.com/Bioconductor/bioc_docker](https://github.com/Bioconductor/bioc_docker)), and available to the community through [DockerHub](https://hub.docker.com/).
Past and current combinations of R and _Bioconductor_ are therefore accessible via a specific docker tag.

@@

Use of images suggests several interesting possibilities for the _Bioconductor_ project.
Images may be valuable in teaching, where participants pull pre-built images to avoid complicated configuration of their own computing environemnts.
An appeal of this over our current approach (providing Amazon Machine Instances for the duration of the course) is the utility of the image to participant after the course is over.
`bioconductor_full` introduces a common system configuration, so it becomes increasingly sensible for _Bioconductor_ to distribute convenient _binary_ packages.
Images also suggest approaches to more advanced computational models.
For instance, we are exploring use of images for [Helm](https://helm.sh/)-orchestrated [Kubernetes](https://kubernetes.io/) clusters on the Google Cloud Platform.
The user interacts with a manager image based on `bioconductor_full`, configured to perform map-reduce style computations via the BiocParallel package communicating with minimally-configured worker images.
A strength of this approach is that the responsibility for complex software configuration (including customized development) is shifted from the user to the experienced _Bioconductor_ core team.

>>>>

_Bioconductor_ ([https://bioconductor.org/](https://bioconductor.org/)) is an open source, open development project for the analysis and comprehension of genomic data \citep{gentleman_bioconductor_2004}.
The project consists of 1741 R software packages as of August 15th 2019, as well as packages containing annotation or experiment data.
_Bioconductor_ has a semiannual release cycle, each release is associated with a particular version of R.
Docker images allow availability of current and past versions of _Bioconductor_ for convenience and reproducibility.
_Bioconductor_'s 'base' docker images are built on top of `rocker/r-ver` and `rocker/rstudio` images.
_Bioconductor_ installs packages based on the R version, and therefore uses `rocker/rstudio` and `rocker/r-ver` version tagging.
_Bioconductor_ selects the desired version of R from Rocker, adds the BiocManager CRAN package for installing appropriate versions of _Bioconductor_ packages, and creates a _Bioconductor_ docker image with an informative tag (R\_version\_Bioc\_version).
The images are summarized on the _Bioconductor_ web site (see [https://bioconductor.org/help/docker/](https://bioconductor.org/help/docker/)), created with `Dockerfile`s maintained on GitHub ([https://github.com/Bioconductor/bioconductor_docker](https://github.com/Bioconductor/bioconductor_docker)), and [distributed through DockerHub](https://hub.docker.com/u/bioconductor).
Past and current combinations of R and _Bioconductor_ are therefore accessible via a specific image tag.
A strength of this approach is that the responsibility for complex software configuration (including customized development) is shifted from the user to the experienced _Bioconductor_ core team.
Images may be valuable in teaching, where participants pull pre-built images to avoid complicated configuration of their own computing environemnts.

@@
